This paper presents an approach to provide strong assurance of the secure
execution of distributed event-driven applications on shared
infrastructures, while relying on a small Trusted Computing Base. We build
upon and extend security primitives provided by Trusted Execution
Environments (TEEs) to guarantee authenticity and integrity properties of
applications, and to secure control of input and output devices.  More
specifically, we guarantee that if an output is produced by the
application, it was allowed to be produced by the application's source code
based on an authentic trace of inputs.

We present an integrated open-source framework to develop, deploy, and use
such applications across heterogeneous TEEs.  Beyond authenticity and
integrity, our framework optionally provides confidentiality and a notion
of availability, and facilitates software development at a high level of
abstraction over the platform-specific TEE layer. We support event-driven
programming to develop distributed enclave applications in Rust and C for
heterogeneous TEE, including Intel SGX, ARM TrustZone and Sancus.

In this article we discuss the workings of our approach, the extensions we made
to the Sancus processor, and the integration of our development model with
commercial TEEs. Our evaluation of security and performance aspects show that
TEEs, together with our programming model, form a basis for powerful security
architectures for dependable systems in domains such as Industrial Control
Systems and the Internet of Things, illustrating our framework's unique
suitability for a broad range of use cases which combine cloud processing,
mobile and edge devices, and lightweight sensing and actuation.

